---
title: "Chapter 7"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r}
library(dplyr)
library(magrittr)
```


# Example 1: Occasionally Dishonest Casino

```{r}
library(HMM)
```


## Initialization
```{r}
# Set the name of states and symbols, and matrices of model parameters
States <- c("Fair", "Loaded")
Symbols <- 1:6
Trans <- matrix(c(0.95, 0.05, 0.1, 0.9), nrow=2, ncol=2, byrow=TRUE)
Emiss <- matrix(c(rep(1/6, 6), c(rep(0.1, 5), 0.5)), nrow=2, ncol=6, byrow=TRUE)
Start <- c(0.5, 0.5)

# Create a HMM
hmm <- initHMM(States, Symbols, startProbs=Start, transProbs=Trans, emissionProbs=Emiss)
hmm
```

## Simulation

```{r}
# Simulate a path of states and sequence of observations
nSim <- 10000
sim <- simHMM(hmm, nSim)
```

```{r}
# Print the first 50 states and observations
sim$states[1:50]
sim$observation[1:50]
```

## Using Observations to get States 

* Assuing previously simulated are the true parameter estimates

### Initialization

```{r}
# Define a function that generates random probability vector
ranProb <- function(n) {
   p <- runif(n)
   return (p / sum(p))
}
  
# Initialize parameter estimates
initStart <- ranProb(2)
initTran <- matrix(c(ranProb(2),ranProb(2)), nrow=2, byrow=TRUE)
initEmiss <- matrix(c(ranProb(6),ranProb(6)), nrow=2, byrow=TRUE)

# Initialize a two-state training HMM
States <- c("St1", "St2")
train <- initHMM(States, Symbols, startProbs=initStart, 
                 transProbs=initTran, emissionProbs=initEmiss)
train
```


### Training

#### Baum-Welch

```{r}
# Train the HMM by Baum-Welch and print the estimated parameters
BWfit <- baumWelch(train, sim$observation, maxIteration=100, 
                   delta=1E-9, pseudoCount=1)
BWfit$hmm
```

* Baum Welch is more accurate

### Viterbi

* Computationally efficient but less accurate

```{r}
# Train the HMM by Viterbi and print the estimated parameters
VTfit <- viterbiTraining(train, sim$observation, maxIteration=1000, 
                         delta=1E-9, pseudoCount=1)
VTfit$hmm

```

## Decoding

```{r}
# Rename the states of the fitted HMM according to emissioni distributions
# Rename the states of the fitted HMM according to emissioni distributions
fit <- initHMM(c("Fair","Loaded"), BWfit$hmm$Symbols, BWfit$hmm$startProbs,
               BWfit$hmm$transProbs, BWfit$hmm$emissionProbs)
```

```{r}
fit
```


### Viterbi


```{r}
# Viterbi decoding using simulated observations and fitted model
vit <- viterbi(fit, sim$observation)
# Print the first 100 predicted states
vit[1:100]
```




### Forward and Backward

```{r}
# Calculate forward and backward probabilities (in log scale)
# using simulated observations and fitted model
fp <- forward(fit, sim$observation)
bp <- backward(fit, sim$observation)
fp[,1:5]
bp[,1:5]
```

## Marginal Probability

```{r}
# Calculate marginal probability (in log scale) of observed sequence
x <- fp[1,nSim]
y <- fp[2,nSim]
if (x>y) mProb <- x+log(1+exp(y-x)) else mProb <- y+log(1+exp(x-y))

# Calculate posterior probabilities of each state for each observation
posterior <- exp((fp + bp) - mProb)

mProb
posterior[,1:5]

# Posterior decoding using simulated observations and fitted model
post <- vector()
for (i in 1:nSim) {
   if (posterior[1,i] > posterior[2,i]) post[i] <- "Fair" 
   else post[i] <- "Loaded"
}
post[1:100]
```

## Simulation Plotting

```{r}

# Plot the simulated (observed) sequence of rolls (first 1000 throws)
nfirst <- 1000
plot(sim$observation[1:nfirst], ylim = c(-9.5, 6), pch = 3, 
     main = "Fair die and loaded die",
     xlab = "Throw no.", ylab = "", bty = "n", yaxt = "n")
axis(2, at = 1:6)

# Compare true path and most probable paths (Viterbi and posterior)
text(0, -1.2, adj = 0, cex = 0.8, col = "black", "True path: green = fair die")
for (i in 1:nfirst) {
   if (sim$states[i] == "Fair")
      rect(i, -1, i + 1, 0, col = "green", border = NA)
   else rect(i, -1, i + 1, 0, col = "red", border = NA)
}
text(0, -3.2, adj = 0, cex = 0.8, col = "black", "Most probable path (Viterbi)")
for (i in 1:nfirst) {
   if (vit[i] == "Fair")
      rect(i, -3, i + 1, -2, col = "green", border = NA)
   else rect(i, -3, i + 1, -2, col = "red", border = NA)
}
text(0, -5.2, adj = 0, cex = 0.8, col = "black", "Most probable path (posterior)")
for (i in 1:nfirst) {
   if (post[i] == "Fair")
      rect(i, -5, i + 1, -4, col = "green", border = NA)
   else rect(i, -5, i + 1, -4, col = "red", border = NA)
}
points(posterior[1, 1:nfirst] - 5, type = "l")

# Show difference between true path and most probable paths
text(0, -7.2, adj = 0, cex = 0.8, col = "black", "Difference (true vs Viterbi)")
differ <- !(sim$states == vit)
for (i in 1:nfirst) {
   if (differ[i])
      rect(i, -7, i + 1, -6, col = rgb(0.3, 0.3, 0.3), border = NA)
   else rect(i, -7, i + 1, -6, col = rgb(0.9, 0.9, 0.9), border = NA)
}
text(0, -9.2, adj = 0, cex = 0.8, col = "black", "Difference (true vs posterior)")
differ <- !(sim$states == post)
for (i in 1:nfirst) {
   if (differ[i])
      rect(i, -9, i + 1, -8, col = rgb(0.3, 0.3, 0.3), border = NA)
   else rect(i, -9, i + 1, -8, col = rgb(0.9, 0.9, 0.9), border = NA)
}






```


# Example 2: Earthquake Time Series

```{r}
library(depmixS4)
```



```{r}
# Set random seed to produce same results (for illustrative purpose)
set.seed(100419)

# Read Earthquakes Count Data
eqdata <- read.csv("/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/Lecture codes/R codes and data for Chapter 7 examples-20200202/earthquakes_data.csv",  header=TRUE, sep=",")

head(eqdata)
```


## Initialization

```{r}
Year <- eqdata[,1]
nEQ <- eqdata[,2]
n <- nrow(eqdata)

# Create a 2 state Poisson-HMM
hmm <- depmix(nEQ~1, nstates=2, family=poisson(), ntimes=n, 
              trstart=runif(4), respstart=runif(2,0,5))
hmm
```

## Training

```{r}
# Fit the model by EM algorithm and returns an object of class depix.fitted
fithmm <- fit(hmm, emcontrol=em.control(maxit=500, tol=1e-8, random.start=TRUE))

# Show convergence message
fithmm

# Print out parameter estimates
summary(fithmm)
```

## Decoding

```{r}
# Determine the viterbi states and posterior probabilities
predStates <- posterior(fithmm)
x <- cbind(Year, nEQ, predStates)

# Create time series plot for the annual number of earthquakes
plot(nEQ~Year, ylim = c(0, 50), xlim = c(1900, 2010), 
     main = "Annual number of earthquakes", xlab = "Year", 
     ylab = "Earthquakes", type="l")

# Display most probable (Viterbi) states in the time series plot
text(1900, 0.5, adj = 0, cex = 0.8, col = "black", 
     "Red: State 1 (rate=26.0)        Green: State 2 (rate=15.4)")
for (i in 1:n) {
   if (predStates[i,1] == 1)
      rect(1899+i, 2, 1900+i, 4, col = "red", border = NA)
   else rect(1899+i, 2, 1900+i, 4, col = "green", border = NA)
}

```




