---
title: "Chapter 5"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r}
library(magrittr)
library(dplyr)
library(locfit)
```


```{r}
lidar <- read.csv("~/MSTAT/STAT6014_Advanced_Stat_Model/Lecture codes/R codes and data for Chapter 5 examples-20200202/lidar.csv")

head(lidar)
```

# Local Polynomial Regression (LIDAR)

## Local Linear Regression, `degree=1`

* smoothing parameter, $\alpha = 0.2$
* 0.2 x 221 = 44 points , k = 44


```{r}
# Construct scatter plot for logratio vs range
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)", 
     pch = 20, cex = 0.7, main="Loess with local linear regression")

# Perform local linear fitting with smoothing parameter 0.2
llfit <- loess(logratio~range, data = lidar, span = 0.2, degree = 1)

# Plot the fitted curve
lines(llfit$x, llfit$fitted, col = "blue", lwd=2)
```


### Predictions

```{r}
# Predictions at range = 500, 550, 600
predict(llfit, newdata = c(500, 550, 600))
```

## Local Quadratic Regression, `degree=2`

```{r}
# Perform local quadratic fitting with smoothing parameter 0.2
lqfit <- loess(logratio~range, data = lidar, span = 0.2, degree = 2)
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)", 
     pch = 20, cex = 0.7, main = "LOESS with local quadratic regression")
lines(lqfit$x, lqfit$fitted, col = "blue", lwd=2)
```


### Predictions

```{r}
# Predictions at range = 500, 550, 600
predict(lqfit, newdata = c(500, 550, 600))
```



## Optimizing Smoothing Parameter, alpha

* using either AIC or GCV


### GCV

```{r}
lqs <- gcvplot(logratio~range, data = lidar, deg = 2,
               alpha = seq(0.1, 0.9, by = 0.005))

# Plot GCV vs smoothing parameter
plot(lqs$alpha, lqs$values, main = "Smoothing parameter Selection GCV",
     xlab = "Smoothing Parameter", ylab = "GCV", pch = 20, cex = 0.7)

# Select smoothing parameter that minimizes GCV
best <- lqs$alpha[which.min(lqs$values)]
legend("topleft", paste0("Selected Smoothing Parameter = ",best))
```

### AIC

```{r}
lqs <- aicplot(logratio~range, data = lidar, deg = 2,
               alpha = seq(0.1, 0.9, by = 0.005))

# Plot GCV vs smoothing parameter
plot(lqs$alpha, lqs$values, main = "Smoothing parameter Selection AIC",
     xlab = "Smoothing Parameter", ylab = "GCV", pch = 20, cex = 0.7)

# Select smoothing parameter that minimizes GCV
best <- lqs$alpha[which.min(lqs$values)]
legend("topleft", paste0("Selected Smoothing Parameter = ",best))
```


### Using Optimized Smoothing Parameter

```{r}
# Perform local quadratic fitting with selected smoothing parameter
lqfit2 <- loess(logratio~range, data = lidar, span = 0.435, degree = 2)
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)", 
     pch = 20, cex = 0.7, main = "LOESS with local quadratic using optimized alpha")
lines(lqfit2$x, lqfit2$fitted, col = "blue", lwd=2)
```

## Diagnostic Plots


```{r}
# Construct residual plots
fv <- lqfit2$fitted
resid <- lqfit2$residuals
layout(matrix(1:4,c(2,2), byrow = TRUE))
# Histogram of Residuals
hist(resid, xlab = "Residuals", prob = TRUE, col = "grey", main = NULL)
curve(dnorm(x, mean = mean(resid), sd = sqrt(var(resid))), add = TRUE,
      col = "darkblue", lwd = 2)

# Boxplot of residuals
boxplot(resid, col = "grey", ylab = "Residuals")

# Q-Q plot of residuals
qqnorm(resid, ylab = "Residuals", pch = 20, main = NULL)
qqline(resid)

# Residuals vs fitted values
plot(fv, resid, pch = 20, xlab = "Fitted Values", ylab = "Residuals")
```
* Residual scatterplot shows higher range at lower values compared to higher values

* Some transformation can stabilize it


# Spline Regression (LIDAR)


## Using Optimized Lambda from GCV

```{r}
# Fit smoothing spline regression with smoothing parameter 
# selected based on GCV
spfit <- smooth.spline(lidar$range, lidar$logratio, cv = FALSE)

# Plot the fitted regression line
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)",
     main = "Cubic smoothing spline fit for LIDAR data",
     pch = 20, cex = 0.7)
lines(spfit, col = "blue", lwd=2)
legend("topright", paste0("lambda = ",round(spfit$lambda,6)))
```

## Varying Values of Lambda

```{r}
# Fit smoothing spline regression with different smoothing parameters
spfit2 <- smooth.spline(lidar$range, lidar$logratio, lambda = 0.00001)
spfit3 <- smooth.spline(lidar$range, lidar$logratio, lambda = 0.0007)
spfit4 <- smooth.spline(lidar$range, lidar$logratio, lambda = 0.05)
spfit5 <- smooth.spline(lidar$range, lidar$logratio, lambda = 3.5)
layout(matrix(1:4,c(2,2), byrow = TRUE))
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)",
     main = paste0("lambda = ", round(spfit2$lambda,6)),
     pch = 20, cex = 0.7)
lines(spfit2, col = "blue", lwd=2)
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)",
     main = paste0("lambda = ", round(spfit3$lambda,6)),
     pch = 20, cex = 0.7)
lines(spfit3, col = "blue", lwd=2)
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)",
     main = paste0("lambda = ", round(spfit4$lambda,6)),
     pch = 20, cex = 0.7)
lines(spfit4, col = "blue", lwd=2)
plot(lidar$range, lidar$logratio, xlab = "Range", ylab = "log(Ratio)",
     main = paste0("lambda = ", round(spfit5$lambda,6)),
     pch = 20, cex = 0.7)
lines(spfit5, col = "blue", lwd=2)
```

