---
title: "Chapter 4"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, message=FALSE}
library(magrittr)
library(dplyr)
library(tidyverse)
library(patchwork)
library(ks)
```

# HSI Dataset

```{r}
hsi <-read.csv(file="/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/data/HSI.csv", header=TRUE,sep=",")
```

* Bandwidth multiplier, `c=1`

```{r}
# Compute kernel density estimates,
# using Quadratic kernel, default bandwidth
kde1 <- density(hsi$Return, kernel = "epanechnikov")

# Construct histogram with binwidth 0.04
hist(hsi$Return, breaks = seq(-0.6, 0.8, 0.04), col = "grey", freq = FALSE,
     main = "Histogram of HSI Returns c = 1", xlab = "HSI Return")
# Add kde to histogram
lines(kde1, col = "blue", lwd = 2)
```

```{r}
kde1
```

* Compare `x` to the density estimates

## Reduce Bandwidth by half `adjust=0.5`

* `h=c\timesbw` 

```{r}
# Compute kernel density estimates,
# using Quadratic kernel, half default bandwidth
kde2 <- density(hsi$Return, adjust = 0.5, kernel = "epanechnikov")
hist(hsi$Return, breaks = seq(-0.6, 0.8, 0.04), ylim = c(0,7), col = "grey", 
     freq = FALSE, main = "Histogram of HSI Returns, c = 0.5", xlab = "HSI Return")
lines(kde2, col = "blue", lwd = 2)
```


## Doubling the bandwidth

```{r}
# Compute kernel density estimates,
# using Quadratic kernel, double default bandwidth
kde3 <- density(hsi$Return, adjust = 2, kernel = "epanechnikov")
hist(hsi$Return, breaks = seq(-0.6, 0.8, 0.04), col = "grey", freq = FALSE, 
     main = "Histogram of HSI Returns,c = 2", xlab = "HSI Return")
lines(kde3, col = "blue", lwd = 2)
```

* `h` and `c` controls the smoothness of the density estimate

* Also known as the smoothing parameter

## Triangular Kernel

```{r}
# Compute kernel density estimates,
# using Trangular kernel, default bandwidth
kde4 <- density(hsi$Return, kernel = "triangular")
hist(hsi$Return, breaks = seq(-0.6, 0.8, 0.04), col = "grey", freq = FALSE, 
     main = "Histogram of HSI Returns", xlab = "HSI Return")
lines(kde4, col = "blue", lwd = 2)
```

## Gaussian Kernel

```{r}
# Compute kernel density estimates,
# using Gaussian kernel, default bandwidth
kde5 <- density(hsi$Return, kernel = "gaussian")
hist(hsi$Return, breaks = seq(-0.6, 0.8, 0.04), col = "grey", freq = FALSE, 
     main = "Histogram of HSI Returns", xlab = "HSI Return")
lines(kde5, col = "blue", lwd = 2)
```

* It seems kernel choice is not that sensitive / crucial compared to **choice of bandwidth**

* Hard to distinguish visually


# Channel Length Data

* Multimodal distribution

```{r}
# Read Channel Length data
cld <-read.csv(file="/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/data/channellength.csv", header=TRUE,sep=",")
```

## Triangular Kernel

```{r}
# Compute kernel density estimates,
# using different kernels, optimal bandwidth 
tkde <- density(cld$length, bw = "SJ-ste", kernel = "triangular")
ekde <- density(cld$length, bw = "SJ-ste", kernel = "epanechnikov")
gkde <- density(cld$length, bw = "SJ-ste", kernel = "gaussian")
```


```{r}
# Construct histogram fitted with kdes
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "Distribution of Channel Lengths", xlab = "Length")
lines(tkde, col = "blue", lwd = 2)
```

## Epanechnikov Kernel

```{r}
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "Distribution of Channel Lengths", xlab = "Length")
lines(ekde, col = "blue", lwd = 2)
```

## Gaussian Kernel

```{r}
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "Distribution of Channel Lengths", xlab = "Length")
lines(gkde, col = "blue", lwd = 2)
```


## Gaussian and SJPI `bw` Estimation

```{r}
# Compute and plot kernel density estimates,
# using gaussian kernel, optimal bandwidth selected by SJPI
kde <- density(cld$length, bw = "SJ-ste", kernel = "gaussian")
plot(kde, col = "blue", lwd = 2, xlab = "Length",
     main = "Kernel Density Estimation for Channel Lengths")
```

## Computing Estimates

```{r}
# Turn computed kde into a function by interpolation
f <- approxfun(kde$x, kde$y)

# Compute kde at x = 0.6, 1.7, 2.3
f(c(0.6, 1.7, 2.3))

```

```{r}
# Integrate kde from 1.0 to 2.0
integrate(f, lower = 1, upper = 2)
```

## Different `bw` Selection Methods

```{r}
# Compute kernel density estimates,
# using gaussian kernel, different bandwidth selectors
kde.snr <- density(cld$length, bw = "nrd0")
kde.srot <- density(cld$length, bw = "nrd")
kde.sjpi <- density(cld$length, bw = "sj-ste")
kde.ucv <- density(cld$length, bw = "ucv")
```

### SNR

```{r}
# Construct histogram fitted with kdes
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0, 1.25), col = "grey", 
     freq = FALSE, main = "KDE (SNR) for Channel Lengths", xlab = "Length")
lines(kde.snr, col = "blue", lwd = 2)
```

### SROT

```{r}
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "KDE (SROT) for Channel Lengths", xlab = "Length")
lines(kde.srot, col = "blue", lwd = 2)
```

### SJPI

```{r}
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "KDE (SJPI) for Channel Lengths", xlab = "Length")
lines(kde.sjpi, col = "blue", lwd = 2)
```

### UCV

```{r}
hist(cld$length, breaks = seq(0, 3, 0.1), ylim = c(0,1.25), col = "grey", 
     freq = FALSE, main = "KDE (UCV) for Channel Lengths", xlab = "Length")
lines(kde.ucv, col = "blue", lwd = 2)
```

* Seems like `SJPI` and `UCV` captures multimodal quite well without oversmoothing

## Confidence Interval

* with `give.Rkern = TRUE`, only returns the R(K) and not fit the KDE
* CI is dont point wise, point by point

```{r}
RK <- density(cld$length, kernel = "gaussian", give.Rkern = TRUE)
kde <- density(cld$length, kernel = "gaussian")

# se is not a scalar, is a vector
se <- sqrt(kde$y * RK / (kde$n * kde$bw))
alpha <- 0.05
zalpha <- qnorm(1- alpha/2)

ciL <- kde$y - zalpha * se
ciU <- kde$y + zalpha * se

```

### C.I Plot
```{r}
plot(kde, ylim = range(ciU), col = "blue", lwd = 2, 
     main = "Kernel Density Estimation for Channel Lengths")
lines(kde$x, ciL, col = "gray", lwd = 2)
lines(kde$x, ciU, col = "gray", lwd = 2)
legend("topright", legend = c("Estimate", "95% CI"), 
       col = c("blue", "gray"), lwd = 2)
```

### Boot Strap Confidence Bands

* Resample based on quantiles

```{r}
# Construct confidence bands of kde by bootstrapping
nrep <- 10000

# Bootstrap starts
bootkde <- replicate(nrep, 
           {
              # Draw bootstrap sample with replacement
              res <- sample(cld$length, replace = TRUE)

              # Compute KDE from resampled data and extract density estimates
              # at same grid points as the original KDE
              density(res, from = min(kde$x), to = max(kde$x), bw = "sj-ste")$y
           })

# Determine the (rowrise) rank of estimates # at each x-coordinate
bootrank <- apply(bootkde, 1, rank)

# Determine the smallest and largest ranks associated with each bootstrap estimate
rank.min <- apply(bootrank, 1, min)
rank.max <- apply(bootrank, 1, max)

# Determine the percentiles of minimum rank and maximum rank
alpha <- 0.05
rankL <- quantile(rank.min, alpha/2)
rankU <- quantile(rank.max, 1-alpha/2)

# Determine the confidence bands
# By construction, at most alpha/2 of the bootstrap estimates have 
# a coordinate with value smaller(larger) than the lower(upper) limit
bootci <- apply(bootkde, 1, quantile, c(rankL/nrep, rankU/nrep))

# Create plot environment
plot(kde, ylim = range(bootci), col = "blue", lwd = 2, 
     main = "Kernel Density Estimation for Channel Lengths")

# Draw the region for confidence bands
polygon(c(kde$x, rev(kde$x)), c(bootci[1,], rev(bootci[2,])),
        col='gray', density = -0.5, border = F)

# Display the lines for KDE and pointwise confidence interval
lines(kde, col = "blue", lwd = 2)
lines(kde$x, ciL, col = "red", lwd = 2, lty = 3)
lines(kde$x, ciU, col = "red", lwd = 2, lty = 3)
legend("topright", legend = c("Estimate", "95% poinwise CI", "95% Simultaneous CI"),
       col = c("blue", "red", "gray"), lty = c(1, 3, NA), lwd = 2, pch = c(NA, NA, 15))
```


# German Credit Scoring

```{r}
gcd <-read.csv(file="/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/data/german_credit.csv", header=TRUE,sep=",")
x <- data.frame(gcd$Age, gcd$Duration)
```

## Multivariate KDE with diagonal bandwidth

```{r}
# Compute bivariate KDEs for (Age, Duration),
# with independent normal reference bandwidth selector
kde.snrd <- kde(x, H = Hns.diag(x))
```

```{r}
kde.snrd$H
```

## Contour Plot

* Peak is where the green is

```{r}
# Contour plot of bivariate KDE for (Age, Duration)
plot(kde.snrd, display = "filled.contour", cont = seq(10, 90, by = 10),
     col = rev(terrain.colors(10, alpha = 0.8)), xlim = c(15, 80), 
     ylim = c(0, 75), main = "Kernel Density for Age and Duration", 
     xlab = "Age", ylab = "Duration")
```

### Other Bandwidth Selections

```{r}
# Compute bivariate KDEs for (Age, Duration),
# with different bandwidth selectors
kde.snr <- kde(x, H = Hns(x))
kde.sjpi <- kde(x, H = Hpi(x))
kde.bcv <- kde(x, H = Hbcv(x))
```

```{r}
layout(matrix(1:4,c(2,2), byrow = TRUE))
plot(kde.snrd, display = "filled.contour", cont = seq(10, 90, by = 10),
     col = rev(terrain.colors(10, alpha = 0.8)), xlim = c(15, 80), 
     ylim = c(0, 75), main = "Kernel Density (SNRD)", 
     xlab = "Age", ylab = "Duration")
plot(kde.snr, display = "filled.contour", cont = seq(10, 90, by = 10),
     col = rev(terrain.colors(10, alpha = 0.8)), xlim = c(15, 80), 
     ylim = c(0, 75), main = "Kernel Density (SNR)", 
     xlab = "Age", ylab = "Duration")
plot(kde.sjpi, display = "filled.contour", cont = seq(10, 90, by = 10),
     col = rev(terrain.colors(10, alpha = 0.8)), xlim = c(15, 80), 
     ylim = c(0, 75), main = "Kernel Density (SJPI)", 
     xlab = "Age", ylab = "Duration")
plot(kde.bcv, display = "filled.contour", cont = seq(10, 90, by = 10),
     col = rev(terrain.colors(10, alpha = 0.8)), xlim = c(15, 80), 
     ylim = c(0, 75), main = "Kernel Density (BCV)", 
     xlab = "Age", ylab = "Duration")
```

* SJPI can capture the bimodal property and detail
* SNRD and SNR oversmoothing