---
title: "Chapter 8"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r}
library(magrittr)
library(dplyr)
library(bnlearn)
```

# Travel Survey  (Domain Knowledge Input)


## Step 1: Initialization
```{r}
# Create a DAG with no arc
variables = c("age","sex","edu","occ","res","trans")
dag <- empty.graph(nodes = variables)

# Add 6 arcs encoding dependencies
dag <- set.arc(dag, from = "age", to = "edu")
dag <- set.arc(dag, from = "sex", to = "edu")
dag <- set.arc(dag, from = "edu", to = "occ")
dag <- set.arc(dag, from = "edu", to = "res")
dag <- set.arc(dag, from = "occ", to = "trans")
dag <- set.arc(dag, from = "res", to = "trans")

# Print the DAG information and plot the DAG
dag
```

## DAG Plot

```{r}
plot(dag)
```

## Model String to Network

* Can do the above but with model string

```{r}
# Build the network directly from the model
dag2 <- model2network("[age][sex][edu|age:sex][occ|edu][res|edu][trans|occ:res]")

# Compare structures of the two networks
compare(dag, dag2)
all.equal(dag, dag2)
```

## D-Separation and Markov Blanket

```{r}
# Check d-seperation of age and occ given edu
dsep(dag, "age", "occ", "edu")

# Check d-seperation of age and trans given res
dsep(dag, "age", "trans", "res")

# Show the Markov blanket of edu
mb(dag, "edu")
```

## Step 2: Define Levels and Probabilities

```{r}
# Define levels for each node
age.lv <- c("young","adult","old")
sex.lv <- c("M","F")
edu.lv <- c("hs","univ")
occ.lv <- c("emp","self")
res.lv <- c("small","big")
trans.lv <- c("car","train","others")

# Create arrays of parameters for each local distribution
age.prob <- array(c(0.30, 0.50, 0.20), dim = 3, dimnames = list(age = age.lv))
sex.prob <- array(c(0.60, 0.40), dim = 2, dimnames = list(sex = sex.lv))
edu.prob <- array(c(0.75, 0.25, 0.64, 0.36, 0.72, 0.28, 
                    0.70, 0.30, 0.88, 0.12, 0.90, 0.10), dim = c(2, 2, 3),
                  dimnames = list(edu = edu.lv, sex = sex.lv, age = age.lv))
occ.prob <- array(c(0.96, 0.04, 0.92, 0.08), dim = c(2, 2),
                  dimnames = list(occ = occ.lv, edu = edu.lv))
res.prob <- array(c(0.25, 0.75, 0.20, 0.80), dim = c(2, 2),
                  dimnames = list(res = res.lv, edu = edu.lv))
trans.prob <- array(c(0.48, 0.42, 0.10, 0.58, 0.24, 0.18,
                      0.56, 0.36, 0.08, 0.70, 0.21, 0.09), dim = c(3, 2, 2),
                    dimnames = list(trans = trans.lv, res = res.lv, occ = occ.lv))
```


## Step 3: Combine and Fit Model

```{r}
# Combine the network structure with local distributions to form a BN
cpt <- list(age = age.prob, sex = sex.prob, edu = edu.prob, occ = occ.prob, 
            res = res.prob, trans = trans.prob)
tsbn <- custom.fit(dag, cpt)
```

```{r}
tsbn
```

* Those with single variables are marginal distributions
* Nodes without any arrows pointing to them will have marginal distributions
* Nodes with arrows pointing towards them will have conditional distributions

## Querying Information

* Want to query $P(edu = univ | trans=car)$

* Uses monte carlo, hence every query gives diff answer 

* Best if provide n

```{r}
# CPQ: calculate P(edu="univ"|trans="car"), default sample size
cpquery(tsbn, event=(edu=="univ"), evidence=(trans=="car"), method="ls")
cpquery(tsbn, event=(edu=="univ"), evidence=(trans=="car"), method="ls")

# CPQ: calculate P(edu="univ"|trans="car"), large sample size
cpquery(tsbn, event=(edu=="univ"), evidence=(trans=="car"), method="ls", n=100000000)
```

# Travel Survey (Learn From Data)

```{r}
tsdata <- read.csv("/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/Lecture codes/R codes and data for Chapter 8 examples-20200202/travel_survey_data.csv")

head(tsdata)
```


## Learn the DAG from Data 

### Grow Shrink Algorithm

```{r}
# Learn the DAG from data using Grow-Shrink algorithm
gsdag <- gs(tsdata, whitelist=NULL, blacklist=NULL)

# Print the learned DAG information and plot the DAG
gsdag

```


```{r}
plot(gsdag)
```


### Hill Climbing Algorithm (Greedy)

```{r}
# Learn the DAG from data using Hill-Climbing algorithm
hcdag <- hc(tsdata, whitelist=NULL, blacklist=NULL)

# Print the learned DAG information and plot the DAG
hcdag

```



```{r}
plot(hcdag)
```


## Parameter Estimation

### GS 

```{r}
# Parameter estimation based on the DAG learned by GS
gsfit <- bn.fit(gsdag, tsdata, method = "mle")

# Print all the parameter estimates
gsfit
```


### HC

```{r}
# Parameter estimation based on the DAG learned by GS
hcfit <- bn.fit(hcdag, tsdata, method = "mle")

# Print all the parameter estimates
hcfit
```


## Querying

```{r}
cpquery(gsfit, event=(edu=="univ"), evidence=(trans=="car"), method="ls")
cpquery(hcfit, event=(edu=="univ"), evidence=(trans=="car"), method="ls")
```


# Hybrid BN

* Contains both continuous and discreet data

```{r}
# Load the hypothetical data set from the bnlearn package
data(clgaussian.test)

head(clgaussian.test)
```


## Training

```{r}
# Learn the DAG from data using Max-Min Hill-Climbing algorithm
mmhcdag <- mmhc(clgaussian.test, whitelist=NULL, blacklist=NULL)

# Print the learned DAG information
mmhcdag
```

```{r}
# Plot the DAG (highlight continuous nodes in red)
plot(mmhcdag, highlight=c("D","E","H","G"), color="red")
```

* Red variables are the continuous variables
* G is a variable that corresponds to conditional distribution given 4 variables

```{r}
# Parameter estimation based on the DAG learned
fit <- bn.fit(mmhcdag, clgaussian.test, method = "mle")

fit
```

* Continuous and discreet will form a normal distribution
* Mean of the normal distribution is a linear regression given the numerical variable

## Querying

```{r}
cpquery(fit, event=(A=="b"), evidence=(D>5), method="ls")
cpquery(fit, event=(A=="b"), evidence=(C=="a"), method="ls")

```



