---
title: "Tutorial 6"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, message=FALSE}
library(HH)
library(nlme)
library(dplyr)
library(lmtest)
library(lme4)
library(glm2)
library(ggplot2)
```

# Awards Data

Demonstrating Multilevel Model

* Classes intoduce randomness


```{r}
awd <- read.csv("/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/Tutorial codes/Tutorial 6/awards.csv")

awd$school <- as.factor(awd$school)
awd$class <- as.factor(awd$class)

awd
```

## Part (a)

```{r}
fitpr <- glm2(award~exam+trt, data = awd, family = "poisson")

summary(fitpr)
```

```{r}
772.90 / 317
```

Shows that it underfits the data. Consider mixed effects.

## Part (b)

Fit Mixed Poisson regression model, with
  * fixed effects: trt, exam
  * random effects: independent intercept and slope of trt within class and school
  
Fitted model is hence

$$log(\hat\mu_{ijk}) = \beta_0+\beta_1t_{ij}+\beta_2x_{ijk}+a_i+b_it_{ij}+c_{ij}+d_{ij}t_{ij}$$

* **Response**: $y_{ijk}$, award earned by the kth student from the jth class of the ith school

* **School effects**: $a_i, b_i$, random intercept and slope of the ith school

* **Class effects**: $c_{ij}, d_{ij}$, random intercept and slope of the jth class in the ith school


```{r}
fit1 <- glmer(award~exam+trt+(1+trt||school/class), data = awd, family = "poisson")

summary(fit1)
```

Fitted model is:

$$log(\hat\mu_{ijk}) = -1.36 + 0.93t{ij} + 0.019x_{ijk} + a_i+b_it_{ij}+c_{ij}+d_{ij}t_{ij} $$

Variance components of class effects:

* $\sigma_c = 0.68$
* $\sigma_d = 6.313\times10^{-5}$

Variance components of school effects:

* $\sigma_a = 0.50$
* $\sigma_b = 4.36\times10^{-5}$


## Part (c)

### Remove Treatment Effects

The small variances of $\sigma_b$ and $\sigma_d$ indicates that the treatment effects within class and school can be removed

```{r}
# Fit Mixed Poisson regression model, with
# fixed effects: trt, exam
# random effects: random intercepts within class and school
fit2 <- glmer(award ~ trt + exam + (1 | school/class),
             family = poisson, data = awd)
summary(fit2)
```

The model fitted is now

$$log(\hat\mu_{ijk})=-1.36+0.93t_{ij}+0.019x_{ijk}+a_i+c_{ij}$$

There is almost no change in parameter estimates


### Random intercept only for class within school 

```{r}
fit3 <- glmer(award ~ trt + exam + (1 | school:class),
             family = poisson, data = awd)
summary(fit3)
```

Fitted model is now:

$$log(\mu_{ijk}) = -1.31+0.83t_{ij}+0.019x_{ijk}+ c_{ij}$$

```{r}
lrtest(fit2, fit3)
```

```{r}
# adjusted p-value due to boundary issue
0.5*0.0357
```

* Less than 0.05 which means the school effects intercept is significant

* The number of awards earned by students are significantly differently among schools.


## Part (d)

```{r}
fixed.effects(fit2)
```

$$log(\hat\mu_{ijk})=-1.36+0.93t_{ij}+0.019x_{ijk}+a_i+c_{ij}$$

```{r}
exp(c(0.93018057, 0.01890311))
```

* on average, students attending ”flipped classroom” are expected to earn 153.5% more awards than students attending traditional classes and with same exam performance

* on average, students with **1 more exam score** are expected to earn 1.9% more awards for the same teaching method.


## Part (e)

for a student from class 6 (trt = 1) in school 2 with exam score 82.7, what is the expected number of awards earned?

```{r}
coef(fit2)$`class:school`[6, ]
```

$$\hat\mu=exp(-1.592174+0.9301806+0.01890311\times82.7)=2.462$$

## Part (f)

For a student in school 2 attended ”flipped classroom” with exam score 82.7, what is the expected number of awards earned?

```{r}
coef(fit2)$school[2, ]
```

$$\hat\mu = exp(-0.6434734+0.9301806+0.01890311\times82.7)=6.358$$

