---
title: "Tutorial 1"
output: 
  html_notebook:
    toc: true
---

# Libraries 

```{r}
library(magrittr)
library(dplyr)
library(tidyverse)
library(psych)
```

# Dplyr 

Main useful functions

* `mutate()`
* `select()`
* `filter()`
* `arrange()`

```{r}
head(mtcars)
```


```{r}
str(mtcars)
```

```{r}
describe(mtcars, skew = F, ranges = F)
```

## Mutate

Adding new variables/columns

```{r}
mtcarsnew1 <- mtcars %>% mutate(kpg = 1.609*mpg) ; str(mtcarsnew1)
```

## Transmute

* Does not return the whole df, but just the constructed column

```{r}
mtcarsnew2 <- mtcars %>% transmute(kpg = 1.609*mpg) ; mtcarsnew2
```

## Select

selects / subsets some columns

```{r}
mtcars.sel1 <- mtcars %>% select(c("vs", "mpg", "hp")) ; mtcars.sel1
```

Select columns that starts with (sub)string
```{r}
mtcars.sel2 <- mtcars %>% select(starts_with('d')) ; mtcars.sel2
```

## Filter

Filter rows with conditions

```{r}
mtcars.fil1 <- mtcars %>% filter(carb==4 & gear==3) ; mtcars.fil1
```

```{r}
mtcars.fil2 <- mtcars %>% filter(hp > mean(hp)); mtcars.fil2
```

## Arrange

For sorting column(s)

```{r}
mtcars %>% arrange(disp)
```

```{r}
mtcars %>% arrange(desc(disp))
```

```{r}
mtcars %>% arrange(cyl, disp)
```

## Group by

```{r}
mtcars %>% group_by(cyl) %>% summarise(disp_mu=mean(disp), hp_mu=mean(hp))
```

### Apply to all 

```{r}
mtcars %>% 
  group_by(cyl) %>% 
    summarise_all(funs(mean = mean(., na.rm = TRUE)))
```

### Apply to some columns

```{r}
mtcars %>% 
  group_by(cyl) %>% 
    summarise_at(vars(mpg, disp, hp), funs(mean = mean(., na.rm = TRUE)))
```

