---
title: "Tutorial 4"
output: 
  html_notebook:
    toc: true
---

# Libraries


```{r}
library(glmulti)
library(glm2)
library(magrittr)
library(tidyverse)
library(dplyr)
library(lmtest)
library(MASS)
library(glue)
```

# Resistivity Data

Using the resistivity dataset , fit a GLIM with the gamma distribution and the log link, with second order interactions.

```{r}
resd <- read.csv("/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/data/resistivity.csv")

resd
```

## Fit Gamma GLIM with Interactions (Full Model)

```{r}
full <- glm2(resistivity~(x1+x2+x3+x4)*(x1+x2+x3+x4), family = Gamma(link = "log"), data = resd)

summary(full)
```

### Find the MLE of Dispersion Parameter

```{r}
summary.full <- summary(full, dispersion = gamma.dispersion(full))
summary.full
```

### Full Model Scaled Deviance

```{r}
scaled_dev_full <- summary.full$deviance /  gamma.dispersion(full) 
scaled_dev_full
```

## Reduced Model using `glmulti`

Find the best sub models of the full model using `glmulti`.

```{r}
all <- glmulti(resistivity~x1+x2+x3+x4, data = resd, family=Gamma(link = "log"), level = 2, method = "h", crit = "aic")
all
```

```{r}
weightable(all)
```


```{r}
all@formulas[[1]]
```

### Fit the Best Reduced Model

```{r}
reduce <- glm2(resistivity ~ 1 + x1 + x2 + x3 + x4 + x3:x1 + x3:x2 + x4:x3, family = Gamma(link = "log"), data = resd)

summary.reduce <- summary(reduce, dispersion = gamma.dispersion(reduce))
summary.reduce
```

### Reduced Model Scaled Deviance

```{r}
scaled_dev_reduced <- summary.reduce$deviance / gamma.dispersion(reduce)
scaled_dev_reduced
```

## LR Test to Determine if Reduced Model Acceptable

Is reduced significantly worse?

```{r}
LRT <- scaled_dev_reduced - scaled_dev_full
DF <- reduce$df.residual - full$df.residual
pvalue <- pchisq(LRT, DF, lower.tail = FALSE)
pvalue
```

Hi p-value indicates is not significantly worse.

### Prediction

Based on the selected model, predict the resistivity when x1 and x2 are turned low while x3 and x4 are turned high.

In short, `x1, x2 = -1` and `x3, x4 = 1`

```{r}
reduce
```

Reduce model is as follow:

$$log(\hat{\mu}) =5.41422 + 0.06127x1 - 0.14957x2 + 0.08994x3     -0.02780x4 -0.03887x1x3 - 0.04407x2x3 - 0.04549x3x4$$

```{r}
log_mu <- predict(reduce,newdata=data.frame(x1=-1,x2=-1,x3=1,x4=1))
mu <- predict(reduce,newdata=data.frame(x1=-1,x2=-1,x3=1,x4=1),type="response")

glue("Log mu is {log_mu}")
glue("Mu is {mu}")
```


