---
title: "Tutorial 9"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r}
library(magrittr)
library(dplyr)
library(mgcv)
```

# Question 1


To develop a statistical model that allows the estimation of body fat for men using only a scale and a measuring tape, a sample of 252 men was surveyed, with their bodyfat (%), age (yr), weight (lbs) and wrist circumference (cm) collected in the dataset
bodyfat.csv.

```{r}
bfd <- read.csv("/Users/jasonchan/MSTAT/STAT6014_Advanced_Stat_Model/Tutorial codes/Tutorial9/bodyfat.csv")

head(bfd)
```

## Part (a)

Fit a nonparametric regression function to predict bodyfat from weight, using smoothing spline.

```{r}
# Fit smoothing spline of bodyfat on weight,
# with smoothing parameter selected based on GCV
spfit <- smooth.spline(bfd$weight, bfd$bodyfat, cv = FALSE)
spfit

```

```{r}
# Plot the fitted regression line
plot(bfd$weight, bfd$bodyfat, pch = 20, cex = 0.8,
     main = "Cubic smoothing spline fit for Bodyfat data",
     xlab = "Weight (lb)", ylab = "Bodyfat (%)")
lines(spfit, col = "blue", lwd=2)
info <- c(paste0("lambda = ",round(spfit$lambda,4)),
          paste0("model df = ",round(spfit$df,2)),
          paste0("RSS = ",round(spfit$pen.crit,2)),
          paste0("GCV = ",round(spfit$cv.crit,2)))
legend(280, 15, info)
```


## Part (b)

Fit an additive model to predict bodyfat from all the other variables.

```{r}
# Fit additive model 
bf.gam1 <- gam(bodyfat ~ s(age) + s(weight) + s(wrist), data = bfd)
summary(bf.gam1)
```

```{r}
# Construct fit plots for Model 1
plot(bf.gam1, pages = 1, se = TRUE, shade = TRUE)
```

* Age seems like a good candidate for a linear relationship

## Part (c)

Refine the model in part (b) if there is any linear relationship revealed.

```{r}
## Q1.c
# Fit partial linear additive model (Model 2),
# with age as linear component
bf.gam2 <- gam(bodyfat ~ age + s(weight) + s(wrist), data = bfd)
summary(bf.gam2)
```

## Deviance Test

```{r}
# Compare Model 2 to Model 1 by Chi-square test
anova(bf.gam2, bf.gam1, test = "Chisq")
```

* Model 2 is not significantly different from model 1 in terms of deviance

* Wrist also seems like a good candidate for a linear relationship

## Model 3, age, wrist as linear component

```{r}
# Fit partial linear additive model (Model 3),
# with age and wrist as linear component
bf.gam3 <- gam(bodyfat ~ age + wrist + s(weight) , data = bfd)
summary(bf.gam3)
```

## Deviance Test of Model 2 and Model 3

```{r}
# Compare Model 3 to Model 2 by Chi-square test
anova(bf.gam3, bf.gam2, test = "Chisq")
```

* Model 3 is significantly different from model 2. Non linearity of wrist cannot be ignored
* Model 2 is chosen as final model



```{r}
# Construct fit plots for Model 2
plot(bf.gam2, pages = 1, se = TRUE, shade = TRUE)
```


## Prediction

```{r}

# Predict the bodyfat of a 50 yrs old man 
# whose weigh 210 lbs and wrist circumference 15 cm
predict(bf.gam2, newdata = data.frame(age = 50, weight = 210, wrist = 15))
```


* Always start by fitting all regressors as additive, then look at summary and partial plot
* Continue refining and checking deviance

Final model is then model 2 and is fitted as:

$$\hat{µ} = 7.2814 + 0.2644 age + s_1(weight) + s_2(wrist)$$

