---
title: "Tutorial 2"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, warning=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(glm2)
library(lmtest)
```

# Logistic Regression

A group of 118 children aged between 3 and 6 were sampled. They were asked whether or not they believed in Santa Claus. The results are tabulated below:

```{r}
tbl <- as.table(rbind(c(30,13,15,5), c(5,10,12,28)))
dimnames(tbl) <- list(believe = c("Yes","No"), 
                      age = c("Age3","Age4","Age5","Age6"))
tbl
```

## Question (a) 

Conduct an independence test between “age” and “believe” using the Pearson χ2 test.

<center>
$H_0:$ Age and Believe are independent
</center>
```{r}
chisq.test(tbl)
```

The χ2 Statistic is 34.227 and p-value < 0.0001. Hence, reject the null hypothesis at a 5% level at significance. 

## Question (b)

Suppose “age” is the explanatory variable and “believe” is the response variable. Suggest a suitable regression model for this dataset.

* Logistic regression model with "believe" as response variable and belive as the categorical variable.

## Question (c)

Run the regression model in R, use Age6 as reference class.

```{r}
df <- data.frame("Age"=c("Age3", "Age4", "Age5", "Age6"), 
                 "Yes"=c(30, 13, 15, 5),
                 "No"=c(5,10,12,28))

df$Age <- relevel(df$Age, ref = "Age6")

df
```


```{r}

fit <- glm2(formula = cbind(Yes, No)~Age, family = "binomial", data = df)

summary(fit)
```

## Question (d)

Write down the fitted regression model and estimate the probabilities for each age group.

* The regression model is as follows:

$$logit(\hat{\pi}) = -1.7228 + 3.5145I_{Age3} + 1.9851I_{Age4} + 1.9459I_{Age5}$$

* The Probabilities for each Age are:

```{r}

coefs <- summary(fit)$coefficients[1:4]

logit6 <- coefs[1]
logit5 <- coefs[1] + coefs[4]
logit4 <- coefs[1] + coefs[3]
logit3 <- coefs[1] + coefs[2]

logit_list <- c(logit6, logit5, logit4, logit3)

sigmoid <- function(x){
  return(exp(x)/(1+exp(x)))
}

probs <- sigmoid(logit_list)
probs
```

<center>

${\pi}_{Age6}$ = `r probs[1]` \
${\pi}_{Age5}$ = `r probs[2]` \
${\pi}_{Age4}$ = `r probs[3]` \
${\pi}_{Age3}$ = `r probs[4]`

</center>

## Question (e)

Test for the global significance of this model


Likelihood ratio test

```{r}
fit0 <- glm2(formula = cbind(Yes, No)~1, family = "binomial", data = df)

lrtest(fit0, fit)
```

LR = 37.672 with df = 3, so that p-value < 0.0001. The model is significant at the 5% level of significance.

## Question (f)

Does part (e) give the same conclusion as part (a)?

Yes

## Question (g)

What is the estimated odds-ratio between a 3-year old and a 6-year old believing in Santa Claus? Is it significant?

```{r}
ci <- exp(cbind(OR=coef(fit), confint(fit)))
ci
```

* Odds ratio is 33.6 and confidence interval well exceeds 1. Hence, children aged 3 are significantly more likely to believe in Santa compared to children with ge 6.


## Question (h)

What is the estimated odds-ratio between a 4-year old and a 5-year old believing in Santa Claus? Is it significant?

```{r}
# making 5 year old the reference group
df$Age <- relevel(df$Age, ref = "Age5")
fit5 <- glm2(formula = cbind(Yes, No)~Age, family = "binomial", data = df)

exp(cbind(OR=coef(fit5), confint(fit5)))
```

## Question (i)

Using the contingency table, compute the sample odds-ratio between a 4-year old and a 5-year old believing in Santa Claus.

```{r}
tbl
```

```{r}
(13/10) / (15/12)
```

It is consistent with the odds ratio computed from the fitted logistic regression model. (The little discrepancy is just due to rounding error.)

## Question (j)

Run the regression model again using “age” as a quantitative variable this time.

```{r}
df2<- data.frame("age" = 3:6,
                     "Yes" = c(30,13,15,5),
                     "No" = c(5,10,12,28))
df2
```

```{r}
fit_q <- glm2(formula = cbind(Yes, No) ~ age, family = "binomial", data=df2)

summary(fit_q)
```

Regression model is as follow:

$$logit(\hat{\pi}) = 4.8076 - 1.0286 \times age$$

## Question (k)

What is the effect on the probability (odds?) of believing in Santa if the child is 1 year older?

```{r}
exp(cbind(OR=coef(fit_q), confint(fit_q)))
```

```{r}
1 - 0.3575017
```

reduction in odds of 64.2% of believing in Santa

# Takeaways

* Pearson Chi-sq test for independence
* Use LR test to test global significance, all regressors vs just B0
* Use of OR exp(beta) and confidence intervals (lower) must be above 1 to have significant impact


